```{r,echo=F}
angle_diff <- function(a, b) {
  abs(((a - b + pi) %% (2*pi)) - pi)
}
while(T){
  u = sample(2:5,2)*sample(c(-1,1),2,T)
  v = sample(2:5,2)*sample(c(-1,1),2,T)
  th1 = atan2(v[1],u[1])
  th2 = atan2(v[2],u[2])
  u1 = u[1]
  v1 = v[1]
  u2 = u[2]
  v2 = v[2]
  u3 = sum(u)
  v3 = sum(v)
  if(angle_diff(th1,th2)>pi/6 && abs(angle_diff(th1,th2)-pi) > pi/6 && min(abs(c(u3,v3)))>0){
    break
  }
}

wu = list()
wu[[1]] = c(-u3,v3)
wu[[2]] = c(u3,-v3)
wu[[3]] = c(-u3,-v3)
wu[[4]] = c(v3,u3)
wu[[5]] = c(-v3,u3)
wu[[6]] = c(v3,-u3)
wu[[7]] = c(-v3,-u3)
wu[[8]] = c(u3+1,v3)
wu[[9]] = c(u3-1,v3)
wu[[10]] = c(u3,v3+1)
wu[[11]] = c(u3,v3-1)

shortc = paste0(u3,"_",v3)
shorts = paste0(u3,"_",v3)
choices = list(c(u3,v3))

while(length(choices)<6){
  i = sample(1:11,1)
  cho = wu[[i]]
  short = paste0(cho[1],"_",cho[2])
  if(!(short %in% shorts)){
    choices[[length(choices)+1]] = cho
    shorts = c(shorts,short)
  }
}

ord = sample(1:6)

choices = choices[ord]
shorts = shorts[ord]

sol = paste0(1*(shorts==shortc),collapse="")
  
```


Question
========

Consider a box sitting on a frictionless surface viewed from above. Two force vectors are shown as red arrows.

```{r,fig.dim=c(4,4),echo=F}
par(mar=c(0,0,0,0))
plot(0,0,"n",xlim=c(-10,10),ylim=c(-10,10),axes=F,ann=F)
for(i in -10:10){
  for(j in -10:10){
    points(i,j,pch=19,cex=0.3)
  }
}
polygon(c(-0.4,-0.4,0.4,0.4),c(0.4,-0.4,-0.4,0.4),col=rgb(0,0,0,0.5))
arrows(0,0,u1,v1,length=0.1,angle=15,2,col="red",lwd=2)
arrows(0,0,u2,v2,length=0.1,angle=15,2,col="red",lwd=2)

```
\

Which of the following options shows the **net** force acting on the box?

```{r,echo=F,fig.dim=c(8,12)}
par(mar=c(0,0,0,0),mfrow=c(3,2))
for(i in 1:6){
  plot(0,0,"n",xlim=c(-10,10),ylim=c(-10,10),axes=F,ann=F)
  for(ii in -10:10){
    for(jj in -10:10){
      points(ii,jj,pch=19,cex=0.3)
    }
  }
  polygon(c(-0.4,-0.4,0.4,0.4),c(0.4,-0.4,-0.4,0.4),col=rgb(0,0,0,0.5))
  arrows(0,0,choices[[i]][1],choices[[i]][2],length=0.1,angle=15,2,col="blue",lwd=2)
  text(0,9.5,LETTERS[i],cex=2)
}

```
\



Answerlist
--------

* A
* B
* C
* D
* E
* F



Solution
========

Use the parallelogram rule.


```{r,fig.dim=c(4,4),echo=F}
par(mar=c(0,0,0,0))
plot(0,0,"n",xlim=c(-10,10),ylim=c(-10,10),axes=F,ann=F)
for(i in -10:10){
  for(j in -10:10){
    points(i,j,pch=19,cex=0.3)
  }
}
polygon(c(-0.4,-0.4,0.4,0.4),c(0.4,-0.4,-0.4,0.4),col=rgb(0,0,0,0.5))
arrows(0,0,u1,v1,length=0.1,angle=15,2,col="red",lwd=2)
arrows(0,0,u2,v2,length=0.1,angle=15,2,col="red",lwd=2)
arrows(u1,v1,u3,v3,length=0.1,angle=15,2,col=rgb(1,0,0,0.7),lwd=2)
arrows(u2,v2,u3,v3,length=0.1,angle=15,2,col=rgb(1,0,0,0.7),lwd=2)
arrows(0,0,u3,v3,length=0.1,angle=15,2,col="blue",lwd=2)

```
\

So the answer is `r LETTERS[which(shorts==shortc)]`.

Meta-information
============
extype: schoice
exsolution: `r sol`
exname: vector_add_2d_graph





