---
title: "Honors Geometry Quiz"
output: pdf_document
params:
  seed: 1
geometry: margin=0.75in
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

set.seed(params$seed)


# ============================================================
# HELPER FUNCTIONS
# ============================================================

# Random nonzero integer
nonzero <- function(a, b) {
  sample(c(a:-1, 1:b), 1)
}


# Format something like:
# 2m - 3
# m + 4
# -m - 2
# 3m
format_linear <- function(a, b, variable = "m") {
  
  if (a == 1) {
    first <- variable
  } else if (a == -1) {
    first <- paste0("-", variable)
  } else {
    first <- paste0(a, variable)
  }
  
  if (b > 0) {
    paste0(first, " + ", b)
  } else if (b < 0) {
    paste0(first, " - ", abs(b))
  } else {
    first
  }
}


# Format ax + by = c nicely
format_equation <- function(a, b, c) {
  
  if (a == 1) {
    xpart <- "x"
  } else if (a == -1) {
    xpart <- "-x"
  } else {
    xpart <- paste0(a, "x")
  }
  
  if (b == 1) {
    ypart <- " + y"
  } else if (b == -1) {
    ypart <- " - y"
  } else if (b > 0) {
    ypart <- paste0(" + ", b, "y")
  } else {
    ypart <- paste0(" - ", abs(b), "y")
  }
  
  paste0(xpart, ypart, " = ", c)
}



# ============================================================
# QUESTION 1
# QUADRILATERAL
# ============================================================

# Approximately half of generated quadrilaterals
# will be equilateral.
is_equilateral <- sample(c(TRUE, FALSE), 1)

Ax <- sample(-4:2, 1)
Ay <- sample(-4:2, 1)

# First side vector
choices <- list(
  c(2, 1),
  c(3, 1),
  c(3, 2),
  c(4, 1),
  c(4, 3),
  c(2, 3)
)

u <- choices[[sample(length(choices), 1)]]

# Sometimes point downward
if (sample(c(TRUE, FALSE), 1)) {
  u[2] <- -u[2]
}

if (is_equilateral) {
  
  # A perpendicular vector of exactly the same length.
  # This generates a square.
  v <- c(-u[2], u[1])
  
  if (sample(c(TRUE, FALSE), 1)) {
    v <- -v
  }
  
} else {
  
  # Start with a perpendicular vector and alter it.
  # This gives a parallelogram with unequal adjacent sides.
  v <- c(-u[2], u[1])
  
  v[1] <- v[1] + sample(c(-2, -1, 1, 2), 1)
  
  # Safety check to guarantee unequal side lengths.
  while (sum(u^2) == sum(v^2) || all(v == 0)) {
    v[1] <- v[1] + sample(c(-1, 1), 1)
  }
}

A <- c(Ax, Ay)
B <- A + u
D <- A + v
C <- B + v



# ============================================================
# QUESTION 2
# RIGHT TRIANGLES
# ============================================================

# Problems with missing hypotenuse
legs <- list(
  c(2, 3),
  c(2, 5),
  c(3, 4),
  c(3, 5),
  c(4, 5),
  c(5, 6),
  c(5, 7),
  c(6, 7),
  c(6, 8),
  c(7, 8),
  c(8, 9),
  c(9, 12)
)

p1 <- legs[[sample(length(legs), 1)]]
p2 <- legs[[sample(length(legs), 1)]]

while (all(p1 == p2)) {
  p2 <- legs[[sample(length(legs), 1)]]
}


# Problems with missing leg
# First number = known leg
# Second number = hypotenuse
leg_missing <- list(
  c(3, 5),
  c(4, 7),
  c(5, 8),
  c(5, 9),
  c(6, 10),
  c(7, 11),
  c(8, 13),
  c(9, 15),
  c(10, 17),
  c(12, 17),
  c(15, 17)
)

p3 <- leg_missing[[sample(length(leg_missing), 1)]]
p4 <- leg_missing[[sample(length(leg_missing), 1)]]

while (all(p3 == p4)) {
  p4 <- leg_missing[[sample(length(leg_missing), 1)]]
}



# ============================================================
# QUESTION 3
# POINTS ON A LINE
# ============================================================

# Randomly use a vertical or horizontal line.
vertical <- sample(c(TRUE, FALSE), 1)

# Perpendicular distance from the point to the line
across <- sample(2:6, 1)

# Distance along the line to each solution
along <- sample(3:8, 1)

# Total distance squared
distance_squared <- across^2 + along^2

if (vertical) {
  
  line_value <- sample(-4:4, 1)
  point_y <- sample(-4:4, 1)
  
  point_x <- line_value +
    sample(c(-across, across), 1)
  
  line_text <- paste0("x = ", line_value)
  
} else {
  
  line_value <- sample(-4:4, 1)
  point_x <- sample(-4:4, 1)
  
  point_y <- line_value +
    sample(c(-across, across), 1)
  
  line_text <- paste0("y = ", line_value)
}



# ============================================================
# QUESTION 4
# DISTANCE TWO WAYS
# ============================================================

Jx <- sample(-6:1, 1)
Jy <- sample(-5:3, 1)

distance_vectors <- list(
  c(3, 4),
  c(4, 5),
  c(5, 6),
  c(5, 8),
  c(6, 7),
  c(7, 9),
  c(8, 11),
  c(9, 12)
)

change <- distance_vectors[[sample(length(distance_vectors), 1)]]

dx <- change[1] * sample(c(-1, 1), 1)
dy <- change[2] * sample(c(-1, 1), 1)

Kx <- Jx + dx
Ky <- Jy + dy



# ============================================================
# QUESTION 5
# SYSTEM OF EQUATIONS
# ============================================================

# Choose the desired solution first.
solution_x <- nonzero(-5, 5)
solution_y <- nonzero(-5, 5)

# Generate coefficients until the equations
# are not parallel.
repeat {
  
  a1 <- nonzero(-4, 4)
  b1 <- nonzero(-4, 4)
  
  a2 <- nonzero(-4, 4)
  b2 <- nonzero(-4, 4)
  
  determinant <- a1 * b2 - a2 * b1
  
  if (determinant != 0) {
    break
  }
}

# Construct the constants so the selected point
# is guaranteed to be the solution.
c1 <- a1 * solution_x + b1 * solution_y
c2 <- a2 * solution_x + b2 * solution_y

equation1 <- format_equation(a1, b1, c1)
equation2 <- format_equation(a2, b2, c2)



# ============================================================
# QUESTION 6
# EQUATION OF A LINE
# ============================================================

x1 <- sample(-6:2, 1)
y1 <- sample(-5:5, 1)

change_x <- nonzero(2, 6)
change_y <- nonzero(2, 7)

x2 <- x1 + change_x
y2 <- y1 + change_y



# ============================================================
# QUESTION 7
# SLOPE INVOLVING m
# ============================================================

# Choose the slope that will appear after simplification.
slope_choices <- list(
  c(1, 2),
  c(2, 3),
  c(3, 2),
  c(3, 4),
  c(-1, 2),
  c(-2, 3)
)

pq <- slope_choices[[sample(length(slope_choices), 1)]]

p <- pq[1]
q <- pq[2]

# Common linear factor connecting the two points
a <- sample(1:3, 1)
b <- sample(1:5, 1)

# First point
xA_m <- sample(1:3, 1)
xA_b <- sample(-4:4, 1)

yA_m <- sample(1:3, 1)
yA_b <- sample(-4:4, 1)

# Construct second point.
#
# Difference in x:
# q(am + b)
#
# Difference in y:
# p(am + b)
#
# So the slope simplifies to p/q.
xB_m <- xA_m + q * a
xB_b <- xA_b + q * b

yB_m <- yA_m + p * a
yB_b <- yA_b + p * b


# Format the four coordinates here.
xA_text <- format_linear(xA_m, xA_b)
yA_text <- format_linear(yA_m, yA_b)

xB_text <- format_linear(xB_m, xB_b)
yB_text <- format_linear(yB_m, yB_b)



# ============================================================
# QUESTION 8
# ALGEBRA
# ============================================================

# Squaring a binomial misconception
shift <- sample(2:9, 1)


# Monic trinomial
r1 <- sample(2:9, 1)
r2 <- sample(2:9, 1)

while (r1 == r2) {
  r2 <- sample(2:9, 1)
}

middle <- r1 + r2
constant <- r1 * r2


# Trinomial with a GCF
gcf <- sample(2:5, 1)

f1 <- sample(1:6, 1)
f2 <- sample(1:6, 1)

quad_a <- gcf
quad_b <- gcf * (f1 + f2)
quad_c <- gcf * f1 * f2

```

**Name:** ____________________________________________

**Version:** `r params$seed`

Show all work and clearly indicate your thinking and process. Keep your work organized. Use exact values unless otherwise indicated.

# 1. Coordinate Quadrilateral

Graph the points:

$A(`r A[1]`, `r A[2]`)$

$B(`r B[1]`, `r B[2]`)$

$C(`r C[1]`, `r C[2]`)$

$D(`r D[1]`, `r D[2]`)$

Is quadrilateral $ABCD$ **equilateral**?

Justify your work and explain clearly.


```{r q1-grid, fig.width=5.5, fig.height=5.5}

all_x <- c(A[1], B[1], C[1], D[1])
all_y <- c(A[2], B[2], C[2], D[2])

xmin <- min(-1, all_x) - 2
xmax <- max(1, all_x) + 2

ymin <- min(-1, all_y) - 2
ymax <- max(1, all_y) + 2

plot(
  NULL,
  xlim = c(xmin, xmax),
  ylim = c(ymin, ymax),
  asp = 1,
  xlab = "x",
  ylab = "y",
  xaxt = "n",
  yaxt = "n"
)

x_ticks <- floor(xmin):ceiling(xmax)
y_ticks <- floor(ymin):ceiling(ymax)

abline(
  v = x_ticks,
  h = y_ticks,
  col = "lightgray",
  lwd = 0.7
)

abline(
  h = 0,
  v = 0,
  col = "black",
  lwd = 1.2
)

axis(
  side = 1,
  at = x_ticks
)

axis(
  side = 2,
  at = y_ticks,
  las = 1
)

box()

```


\newpage

# 2. Right Triangles

Find the missing side of each **right triangle**.

The third side listed is always the hypotenuse.

Give each answer in **simplified radical form** when necessary. Show your work.


### a.

`r p1[1]`, `r p1[2]`, ________


\vspace{0.9in}


### b.

`r p2[1]`, `r p2[2]`, ________


\vspace{0.9in}


### c.

`r p3[1]`, ________, `r p3[2]`


\vspace{0.9in}


### d.

________, `r p4[1]`, `r p4[2]`


\newpage

# 3. Points on a Line

Find **both points** on the line

$$
`r line_text`
$$

that are

$$
\sqrt{`r distance_squared`}
$$

units from the point

$$
(`r point_x`, `r point_y`).
$$

Draw a diagram and show your work.


\vspace{1in}


**First point:** ____________________________________


\vspace{0.3in}


**Second point:** ___________________________________


\newpage

# 4. Distance Two Ways

Given the points

$$
J(`r Jx`, `r Jy`)
$$

and

$$
K(`r Kx`, `r Ky`),
$$

answer both questions below.


## a. Pythagorean Theorem

Find the length of segment $JK$ using the **Pythagorean Theorem**.


\vspace{1.5in}


## b. Distance Formula

Find the length of segment $JK$ using the **Distance Formula**.


\vspace{1.5in}


\newpage

# 5. Intersection of Two Lines

Given

$$
`r equation1`
$$

and

$$
`r equation2`,
$$

solve for their intersection.

Give your answer in **exact values** and show your work.


\vspace{2in}


**Intersection:** ___________________________________


\newpage

# 6. Equation of a Line

Given the points

$$
(`r x1`, `r y1`)
$$

and

$$
(`r x2`, `r y2`),
$$

find the equation of the line.

Show enough work to justify your answer.


\vspace{2.25in}


**Equation:** _______________________________________


\newpage

# 7. Slope with Variables

Given the points

$$
(`r xA_text`,\ `r yA_text`)
$$

and

$$
(`r xB_text`,\ `r yB_text`),
$$

find the slope of the line.

Simplify completely and show your algebra.


\vspace{2.25in}


**Slope:** __________________________________________


\newpage

# 8. Algebra


## a. Squaring a Binomial

Does

$$
(x + `r shift`)^2 = x^2 + `r shift^2`
$$

for all values of $x$?

Explain.


\vspace{1.25in}


## b. Factor

Factor completely:

$$
y^2 + `r middle`y + `r constant`
$$


\vspace{1.25in}


## c. Factor Completely

Factor completely:

$$
`r quad_a`v^2 + `r quad_b`v + `r quad_c`
$$

\vspace{1.25in}

\newpage



# 9. Important Question

What is your favorite class other than Geometry?

Why?


\vspace{2in}