---
output: pdf_document
---

**Chapter 3 Practice Quiz**   \hspace{2in} Name: \hrule


**1.** Consider the table below that lists positions and times for a cyclist biking East along the Manhan Rail Trail.

\begin{center}
\begin{tabular}{|c|c|c|c|c|} \hline
$t~$ time (s) & 9 & 11 & 59 & 61 \\ \hline
$x~$ position (m) & 51 & 63 & 467 & 489 \\ \hline
\end{tabular}
\end{center}

**a.** Estimate the velocity of the cyclist around $t=10$ seconds.

\vfill

**b.** Estimate the velocity of the cyclist around $t=60$ seconds.

\vfill

**c.** Estimate the acceleration of the cyclist from $t=10$ until $t=60$.

\vfill


**2.** A baseball is thrown upward at a speed of 40 meters per second. The acceleration due to gravity is $-10 ~\mathrm{m/s^2}$.

**a.** What is the velocity (speed and direction) of the baseball after 2 seconds?

\vfill

**b.** What is the velocity (speed and direction) of the baseball after 7 seconds?

\vfill

**c.** How long does it take for the baseball to switch directions from going up to going down?

\vfill


\newpage

**3.** Consider the **position** versus time graph shown below. Let positive indicate North and negative indicate South.

```{r,echo=F,fig.dim=c(6.5,4)}
par(mar=c(5,5,3,0))
ts = c(0,1,3,9,10)
xs = c(0,4,3,-3,-2)
plot(ts,xs,pch=19,xlim=c(0,10),ylim=c(-5,5),xlab="time (s)",ylab="position (m)",main="Position vs. Time of Object moving North and South")
for(i in 0:10){
  lines(c(i,i),c(-5,5),lwd=0.3)
}
for(i in -5:5){
  lines(c(0,10),c(i,i),lwd=0.3)
}
abline(h=0,lwd=1.5)
lines(ts,xs,lwd=3)
```

**a.** When $t=4$, what is the position of the object? \vfill

**b.** When $t=4$, what is the velocity of the object? (Remember we need speed AND direction.) \vfill

**c.** During which time interval(s) is the object further North than at $t=0$? \vfill

**d.** During which time interval(s) is the object moving North? \vfill


\newpage



**4.** Consider the **velocity** versus time graph shown below. Let positive indicate North and negative indicate South.

```{r,echo=F,fig.dim=c(6.5,4)}
par(mar=c(5,5,3,0))
ts = c(0,4,5,8,10)
xs = c(1,5,0,-3,-2)
plot(ts,xs,pch=19,xlim=c(0,10),ylim=c(-5,5),xlab="time (s)",ylab="velocity (m/s)",main="Velocity vs. Time of Object moving North and South")
for(i in 0:10){
  lines(c(i,i),c(-5,5),lwd=0.3)
}
for(i in -5:5){
  lines(c(0,10),c(i,i),lwd=0.3)
}
abline(h=0,lwd=1.5)
lines(ts,xs,lwd=3)
```


**a.** What is the velocity at $t=7$ s? \vfill

**b.** What is the acceleration at $t=7$ s? \vfill

**c.** When does the object change the direction it is heading? \vfill

**d.** During which intervals is the object slowing down (speed getting closer to 0)? \vfill

\newpage


**5.** A kayaker paddles at a speed of 1.6 m/s relative to water. A river is flowing East at a speed of 1.2 m/s.

**a.** Relative to ground, what is the velocity of the kayaker when the kayak points East? (Remember we need speed and direction for velocity.) Please draw a sketch to indicate the vector addition and direction of the resultant. \vfill

**b.** Relative to ground, what is the velocity of the kayaker when the kayak points West? Please draw a sketch to indicate the vector addition and direction of the resultant. \vfill

**c.** Relative to ground, what is the velocity of the kayaker when the kayak points North? Please draw a sketch to indicate the vector addition and direction of the resultant. \vfill







